### 2020-11-16 paper/results/descriptive_teacher.tex
commit changes to saeb exam scores

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Mon, Nov 09, 2020 - 12:09:20 PM
\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
Number of teachers per school & 72,642 & 4.108 & 4.335 & 1 & 86 \\ 
Number of classes per school & 72,642 & 4.126 & 4.417 & 1 & 173 \\ 
Proportion of teachers w/ higher education & 72,642 & 0.729 & 0.377 & 0 & 1 \\ 
Average age & 72,642 & 0.811 & 0.283 & 0.000 & 1.000 \\ 
Average nunber of classes per teacher & 72,642 & 40.008 & 7.251 & 16.000 & 75.000 \\ 
mean\_n\_class\_per\_teacher & 72,642 & 1.025 & 0.248 & 0 & 7 \\ 
\hline \\[-1.8ex] 
\end{tabular}