### 2020-11-16 paper/results/descriptive_teacher.tex
fix descriptive stats for teacher

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Mon, Nov 16, 2020 - 10:49:28 AM
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
Number of teachers per school & 4.108 & 4.335 & 1 & 86 \\ 
Number of classes per school & 4.126 & 4.417 & 1 & 173 \\ 
Proportion of teachers w/ higher education & 0.729 & 0.377 & 0 & 1 \\ 
Average age & 0.811 & 0.283 & 0.000 & 1.000 \\ 
Average number of classes per teacher & 40.008 & 7.251 & 16.000 & 75.000 \\ 
mean\_n\_class\_per\_teacher & 1.025 & 0.248 & 0 & 7 \\ 
\hline \\[-1.8ex] 
\end{tabular}